---
output:
  word_document: default
  html_document: default
---
## R Markdown---
title: "Final Project: Milestone 5 (Summary of Individual Analysis)"
author: "Tyler Boudart"
date: "5/29/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Add code to upload file in cell below.
```{r LoadData, echo=FALSE, warning=FALSE, message = FALSE}

```

```{r librarys, echo=FALSE, warning=FALSE, message = FALSE}
library(corrplot)
library(dplyr)
library(psych)
library(Amelia)
library(leaps)
library(car)
library(glmnet)
library(caret)
library(vcd)
library(ca)
library(stringr)
library(tidyr)
```

Remove index/ID number and industry.
```{r dataClean, echo=FALSE, warning=FALSE, message = FALSE}
technologyData <- technologyData %>% select(-c(1, 22)) %>% na.omit(technologyData)
```

### Test correlation

```{r corrTest, echo=FALSE, warning=FALSE, message = FALSE}
corrTest <- corr.test(technologyData, adjust="none")
corrSig <- ifelse(corrTest$p < .05, corrTest$r, 0)
corrSig
```


```{r corrTestCount, warning=FALSE, message = FALSE}
corrSigCount <- ifelse(corrTest$p < .05, TRUE, FALSE)
colSums(corrSigCount)-1
```

### Data transformations

```{r dropPrice, echo=FALSE}
techNumeric <- technologyData
```

```{r transformations, echo=FALSE}
techNumeric$assetsLN <- log(technologyData$assets + .00001 - min(technologyData$assets))
techNumeric$gpLN <- log(technologyData$gp + .00001 - min(technologyData$gp))
techNumeric$intexpLN <- log(technologyData$intexp + .00001 - min(technologyData$intexp))
techNumeric$liabilitiesLN <- log(technologyData$liabilities + .00001 - min(technologyData$liabilities))
techNumeric$opexLN <- log(technologyData$opex + .00001 - min(technologyData$opex))
techNumeric$ppnenetLN <- log(technologyData$ppnenet + .00001 - min(technologyData$ppnenet))
techNumeric$rndLN <- log(technologyData$rnd + .00001 - min(technologyData$rnd))
techNumeric$investmentsLN <- log(technologyData$investments + .00001 - min(technologyData$investments))
techNumeric$debtLN <- log(technologyData$debt + .00001 - min(technologyData$debt))
techNumeric$ebitLN <- log(technologyData$ebit + .00001 - min(technologyData$ebit))
techNumeric$ebitdaLN <- log(technologyData$ebitda + .00001 - min(technologyData$ebitda))
techNumeric$ebtLN <- log(technologyData$ebt + .00001 - min(technologyData$ebt))
techNumeric$ncfdivLN <- log(technologyData$ncfdiv + .00001 - min(technologyData$ncfdiv))
techNumeric$marketcapLN <- log(technologyData$marketcap + .00001 - min(technologyData$marketcap))
techNumeric$ncfoLN <- log(technologyData$ncfo + .00001 - min(technologyData$ncfo))
techNumeric$netincLN <- log(technologyData$netinc + .00001 - min(technologyData$netinc))
techNumeric$revenueLN <- log(technologyData$revenue + .00001 - min(technologyData$revenue))
techNumeric$equityLN <- log(technologyData$equity + .00001 - min(technologyData$equity))
```

```{r transformations2, echo=FALSE}
techNumeric$avgTotLiabLN <-  log(techNumeric$avgTotLiab + 0.00001 - min(techNumeric$avgTotLiab))
techNumeric$EBITmarginLN <-  log(techNumeric$EBITmargin + 0.00001 - min(techNumeric$EBITmargin))
techNumeric$ROALN <-  log(techNumeric$ROA + 0.00001 - min(techNumeric$ROA))
techNumeric$debtToAssetsLN <-  log(techNumeric$debtToAssets + 0.00001 - min(techNumeric$debtToAssets))
techNumeric$interestCoverageRatioLN <-  log(techNumeric$interestCoverageRatio + 0.00001 - min(techNumeric$interestCoverageRatio))
techNumeric$sustGrowthRtLN <-  log(techNumeric$sustGrowthRt + 0.00001 - min(techNumeric$sustGrowthRt))
techNumeric$avgTotEquityLN <-  log(techNumeric$avgTotEquity + 0.00001 - min(techNumeric$avgTotEquity))
techNumeric$interestBurdenLN <-  log(techNumeric$interestBurden + 0.00001 - min(techNumeric$interestBurden))
techNumeric$assetTurnoverLN <-  log(techNumeric$assetTurnover + 0.00001 - min(techNumeric$assetTurnover))
techNumeric$ROELN <-  log(techNumeric$ROE + 0.00001 - min(techNumeric$ROE))
techNumeric$debtToEquityLN <-  log(techNumeric$debtToEquity + 0.00001 - min(techNumeric$debtToEquity))
techNumeric$plowBackRtLN <-  log(techNumeric$plowBackRt + 0.00001 - min(techNumeric$plowBackRt))
techNumeric$avgTotAssetsLN <-  log(techNumeric$avgTotAssets + 0.00001 - min(techNumeric$avgTotAssets))
techNumeric$taxBurdenLN <-  log(techNumeric$taxBurden + 0.00001 - min(techNumeric$taxBurden))
techNumeric$profitMarginLN <-  log(techNumeric$profitMargin + 0.00001 - min(techNumeric$profitMargin))
techNumeric$leverageRatioLN <-  log(techNumeric$leverageRatio + 0.00001 - min(techNumeric$leverageRatio))
techNumeric$debtToCapitalLN <-  log(techNumeric$debtToCapital + 0.00001 - min(techNumeric$debtToCapital))
techNumeric$divPayRtLN <-  log(techNumeric$divPayRt + 0.00001 - min(techNumeric$divPayRt))

```

### Explore some relationshops with log market capitalization

```{r plotsv1, warning=FALSE, message = FALSE}
plot(techNumeric$avgTotAssets,techNumeric$marketcapLN)
```

```{r plots2, warning=FALSE, message = FALSE}
plot(techNumeric$assetsLN,techNumeric$marketcapLN)
```

```{r plots2, warning=FALSE, message = FALSE}
plot(techNumeric$assets,techNumeric$marketcap)
```

### Fit OLS and explore residuals

```{r marketCapLMPreliminary}
markCapLmModel <- lm(marketcapLN ~ assets + gp + intexp +  liabilities + opex + ppnenet + rnd, data=techNumeric)
summary(markCapLmModel)
vif(markCapLmModel)
sqrt(mean(markCapLmModel$residuals^2))
```


```{r plotResidual}
plot(markCapLmModel)
```

multiplicative errors.... need to transfer Y with Natural Log 



```{r logMarketCapLog}
markCapLmModelLog <- lm(log(marketcap) ~ assetsLN + investmentsLN, data=techNumeric)
summary(markCapLmModelLog)
vif(markCapLmModelLog)
sqrt(mean(markCapLmModelLog$residuals^2))
```

```{r logMarketCapLMTest}
plot(markCapLmModelLog)
```


```{r logMarketCapLMTrans2}
markCapLmModelLN <- lm(marketcapLN ~ assetsLN + rndLN + investmentsLN, data=techNumeric)
summary(markCapLmModelLN)
vif(markCapLmModelLN)
sqrt(mean(markCapLmModelLN$residuals^2))
```

```{r logMarketCapLMTransPlot}
plot(markCapLmModelLN)
```


# Split data into training and testing sets (75%/25%). 

```{r dataFramesSplit2}
set.seed(210)
tIndex <- createDataPartition(techNumeric[,11], p = .75,
                                  list = FALSE,
                                  times = 1)
train <- techNumeric[tIndex,]
test <- techNumeric[-tIndex,]

xTrain <- as.matrix(train[,c(1:10, 12:51, 53:74)])
yTrain <- as.matrix(train[,52])
xTest <- as.matrix(test[,c(1:10, 12:51, 53:74)])
yTest <- as.matrix(test[,52])
```

### Lasso regression

```{r marketCapLasso}
LassoModel <- cv.glmnet(xTrain, yTrain, alpha = 1)
plot(LassoModel)
```


```{r marketCapLassov2}
markCapLassoModel <- glmnet(xTrain, yTrain, alpha = 1)
markCapLassoModel
```

```{r marketCapLassoCoef}
coef(markCapLassoModel, s = 0.0) 
```

```{r marketCapLassoPredict}
lassoPredTrain <- predict(markCapLassoModel, xTrain, s = 0.225) 
rmseLassoPredTrain <- sqrt(mean((lassoPredTrain - yTrain)^2)) 
rmseLassoPredTrain

lassoPredTest <- predict(markCapLassoModel, xTest, s = 0.225) 
rmseLassoTest <- sqrt(mean((lassoPredTest - yTest)^2)) 
rmseLassoTest


percentChange <- round(((rmseLassoTest/rmseLassoPredTrain)-1)*100, digits = 2)
percentChange
```

```{r marketCapLassoResidualPlot}
residual = lassoPredTest - yTest
plot(lassoPredTest, residual)
```

```{r marketCapLassoResidualQQPlot}
qqnorm(residual, pch = 1, frame = FALSE)
qqline(residual, col = "steelblue", lwd = 2)
```

```{r marketCapLassoResidualStdResidualPlot}
residualStd = residual/sd(residual)
plot(lassoPredTest, residualStd)
```

```{r marketCapLassoResidualStdResidualPercent}
x = residualStd[ residualStd > -3 & residualStd < 3]
length(x)/length(residualStd)
```

### Relaxed Lasso Regression


```{r marketCapLassoRelaxed}
LassoRelaxModel <- cv.glmnet(xTrain, yTrain, alpha = 1, relax = TRUE)
plot(LassoRelaxModel)
```


```{r marketCapLassoRelaxv2}
markCapLassoRelaxModel <- glmnet(xTrain, yTrain, alpha = 1, relax = TRUE)
markCapLassoRelaxModel
```

```{r marketCapLassoRelaxCoef}
coef(markCapLassoRelaxModel, s = 0.225, gamma = 0.0) 
```


```{r marketCapLassoRelaxPredict}
lassoRelaxPredTrain <- predict(markCapLassoRelaxModel, xTrain, s = 0.225, gamma = 0.0) 
rmseLassoRelaxPredTrain <- sqrt(mean((lassoRelaxPredTrain - yTrain)^2)) 
rmseLassoRelaxPredTrain

lassoRelaxPredTest <- predict(markCapLassoRelaxModel, xTest, s = 0.225, gamma = 0.0)  
rmseLassoRelaxTest <- sqrt(mean((lassoRelaxPredTest - yTest)^2)) 
rmseLassoRelaxTest


percentChange <- round(((rmseLassoRelaxTest/rmseLassoRelaxPredTrain)-1)*100, digits = 2)
percentChange
```

cor(tech)

```{r marketCapLassoRelaxResidualPlot}
residual = lassoRelaxPredTest - yTest
plot(lassoRelaxPredTest, residual)
```

```{r marketCapLassoRelaxResidualQQPlot}
qqnorm(residual, pch = 1, frame = FALSE)
qqline(residual, col = "steelblue", lwd = 2)
```

### Test Lasso select feautes with OLS model


```{r logMarketCapLMTrans}
markCapLmModelLN2 <- lm(marketcapLN ~ + assetsLN + investmentsLN, data=techNumeric)
summary(markCapLmModelLN2)
vif(markCapLmModelLN2)
sqrt(mean(markCapLmModelLN2$residuals^2))
```

```{r logMarketCapLMTransLNPlot}
plot(markCapLmModelLN2)
```

### Elastic Net Models

#### alpha = .75

```{r marketCapNet.75}
markCapENet75Model <- glmnet(xTrain, yTrain, alpha = .75)
markCapENet75Model
```


```{r marketCapNet75Coef}
coef(markCapENet75Model, s = 0.249)
```

```{r marketCapNet.75Predict}
Net75PredTrain <- predict(markCapENet75Model, xTrain, s = 0.249) 
rmseNet75PredTrain <- sqrt(mean((Net75PredTrain - yTrain)^2)) 
rmseNet75PredTrain

Net75PredTest <- predict(markCapENet75Model, xTest, s = 0.249) 
rmseNet75Test <- sqrt(mean((Net75PredTest - yTest)^2)) 
rmseNet75Test


percentChange <- round(((rmseNet75Test/rmseNet75PredTrain)-1)*100, digits = 2)
percentChange
```
R-square  = 78.99 per table 

#### alpha = .5


```{r marketCapNet.5}
markCapENet50Model <- glmnet(xTrain, yTrain, alpha = .5)
markCapENet50Model
```

```{r marketCapNet50Coef}
coef(markCapENet50Model, s = 0.4093)
```

```{r marketCapNet50Predict}
Net50PredTrain <- predict(markCapENet50Model, xTrain, s = 0.4093) 
rmseNet50PredTrain <- sqrt(mean((Net50PredTrain - yTrain)^2)) 
rmseNet50PredTrain

Net50PredTest <- predict(markCapENet50Model, xTest, s = 0.4093) 
rmseNet50Test <- sqrt(mean((Net50PredTest - yTest)^2)) 
rmseNet50Test


percentChange <- round(((rmseNet50Test/rmseNet50PredTrain)-1)*100, digits = 2)
percentChange
```
R-square  = 78.46 per table 

#### alpha = .85

```{r marketCapNet.85}
markCapENet85Model <- glmnet(xTrain, yTrain, alpha = .85)
markCapENet85Model
```

```{r marketCapNet85Coef}
coef(markCapENet85Model, s = 0.34950)
```

```{r marketCapNet85Predict}
Net85PredTrain <- predict(markCapENet85Model, xTrain, s = 0.3495) 
rmseNet85PredTrain <- sqrt(mean((Net85PredTrain - yTrain)^2)) 
rmseNet85PredTrain

Net85PredTest <- predict(markCapENet85Model, xTest, s = 0.3495) 
rmseNet85Test <- sqrt(mean((Net85PredTest - yTest)^2)) 
rmseNet85Test


percentChange <- round(((rmseNet85Test/rmseNet85PredTrain)-1)*100, digits = 2)
percentChange
```

Rsquare = 78.21

### Ridge Models

#### Ridge model using all predictors


```{r marketCapRidge}
RidgeModel <- cv.glmnet(xTrain, yTrain, alpha = 0)
plot(RidgeModel)
```


```{r marketCapRidgev2}
markCapRidgeModel <- glmnet(xTrain, yTrain, alpha = 0)
markCapRidgeModel
```


```{r marketRidgeCoef}
coef(markCapRidgeModel, s = RidgeModel$lambda.1se)
```

```{r marketRidgePredict}
RidgePredTrain <- predict(markCapRidgeModel, xTrain, s = RidgeModel$lambda.1se) 
rmseRidgePredTrain <- sqrt(mean((RidgePredTrain - yTrain)^2)) 
rmseRidgePredTrain

RidgePredTest <- predict(markCapRidgeModel, xTest, s = RidgeModel$lambda.1se) 
rmseRidgeTest <- sqrt(mean((RidgePredTest - yTest)^2)) 
rmseRidgeTest


percentChange <- round(((rmseRidgeTest/rmseRidgePredTrain)-1)*100, digits = 2)
percentChange
```

R-square around 77.08

#### Ridge model using three Lasso selected predictors


```{r dataFramesRidge3, echo=FALSE}
xTrain3 <- as.matrix(train[,c("assetsLN", "avgTotAssetsLN", "investmentsLN")])
xTest3 <- as.matrix(test[,c("assetsLN", "avgTotAssetsLN", "investmentsLN")])

```


```{r marketCapRidge3}
RidgeModel3 <- cv.glmnet(xTrain3, yTrain, alpha = 0)
plot(RidgeModel3)
```


```{r marketCapRidge3v2}
markCapRidgeModel3 <- glmnet(xTrain3, yTrain, alpha = 0)
markCapRidgeModel3
```

```{r marketCapRidge3Coef}
coef(markCapRidgeModel3, s = RidgeModel3$lambda.1se)
```


```{r marketRidge3Predict}
Ridge3PredTrain <- predict(markCapRidgeModel3, xTrain3, s = RidgeModel3$lambda.1se) 
rmseRidge3PredTrain <- sqrt(mean((Ridge3PredTrain - yTrain)^2)) 
rmseRidge3PredTrain

Ridge3PredTest <- predict(markCapRidgeModel3, xTest3, s = RidgeModel3$lambda.1se) 
rmseRidge3Test <- sqrt(mean((Ridge3PredTest - yTest)^2)) 
rmseRidge3Test


percentChange <- round(((rmseRidge3Test/rmseRidge3PredTrain)-1)*100, digits = 2)
percentChange
```

R-square 79.59


```{r marketCapRidge3ResidualPlot, echo=FALSE}
residual = Ridge3PredTest - yTest
plot(Ridge3PredTest, residual)
```

```{r marketCapRidge3ResidualQQPlot}
qqnorm(residual, pch = 1, frame = FALSE)
qqline(residual, col = "steelblue", lwd = 2)
```

```{r marketCapRidge3ResidualStdResidualPlot}
residualStd = residual/sd(residual)
plot(Ridge3PredTest, residualStd)
```

```{r marketCapRidge3ResidualStdResidualPercent}
x = residualStd[ residualStd > -3 & residualStd < 3]
length(x)/length(residualStd)
```



#### Ridge model using two Lasso selected predictors



```{r dataFramesRidge2, echo=FALSE}
xTrain2 <- as.matrix(train[,c("assetsLN", "investmentsLN")])
xTest2 <- as.matrix(test[,c("assetsLN", "investmentsLN")])

```

```{r marketCapRidge2}
RidgeModel2 <- cv.glmnet(xTrain2, yTrain, alpha = 0)
plot(RidgeModel2)
```


```{r marketCapRidge2v2, echo=FALSE}
markCapRidgeModel2 <- glmnet(xTrain2, yTrain, alpha = 0)
markCapRidgeModel2
```

```{r marketCapRidge2Coef}
coef(markCapRidgeModel2, s = 0)
```


```{r marketRidge2Predict}
Ridge2PredTrain <- predict(markCapRidgeModel2, xTrain2, s = 0) 
rmseRidge2PredTrain <- sqrt(mean((Ridge2PredTrain - yTrain)^2)) 
rmseRidge2PredTrain

Ridge2PredTest <- predict(markCapRidgeModel2, xTest2, s = 0) 
rmseRidge2Test <- sqrt(mean((Ridge2PredTest - yTest)^2)) 
rmseRidge2Test


percentChange <- round(((rmseRidge2Test/rmseRidge2PredTrain)-1)*100, digits = 2)
percentChange
```

R-square 79.31

```{r marketCapRidgeResidualPlot}
residual = Ridge2PredTest - yTest
plot(Ridge2PredTest, residual)
```

```{r marketCapRidge2ResidualQQPlot}
qqnorm(residual, pch = 1, frame = FALSE)
qqline(residual, col = "steelblue", lwd = 2)
```

```{r marketCapRidgeResidual2StdResidualPlot}
residualStd = residual/sd(residual)
plot(Ridge3PredTest, residualStd)
```

```{r marketCapRidgeResidual2StdResidualPercent}
x = residualStd[ residualStd > -3 & residualStd < 3]
length(x)/length(residualStd)
```

### More Elastic Net Models, alpha = .90




```{r marketCapNet90}
markCapENet90Model <- glmnet(xTrain, yTrain, alpha = .90)
markCapENet90Model
```

```{r marketCapNet90Coef}
coef(markCapENet90Model, s = 0.3006)
```

```{r marketCapNet90Predict}
Net90PredTrain <- predict(markCapENet90Model, xTrain, s = 0.3006) 
rmseNet90PredTrain <- sqrt(mean((Net90PredTrain - yTrain)^2)) 
rmseNet90PredTrain

Net90PredTest <- predict(markCapENet90Model, xTest, s = 0.3006) 
rmseNet90Test <- sqrt(mean((Net90PredTest - yTest)^2)) 
rmseNet90Test


percentChange <- round(((rmseNet90Test/rmseNet90PredTrain)-1)*100, digits = 2)
percentChange
```

R-squared = 78.21

### More Elastic Net Models, alpha = .95

```{r marketCapNet95}
markCapENet95Model <- glmnet(xTrain, yTrain, alpha = .95)
markCapENet95Model
```

```{r marketCapNet95Coef}
coef(markCapENet95Model, s = 0.2595)
```

```{r marketCapNet95Predict}
Net95PredTrain <- predict(markCapENet95Model, xTrain, s = 0.2595) 
rmseNet95PredTrain <- sqrt(mean((Net95PredTrain - yTrain)^2)) 
rmseNet95PredTrain

Net95PredTest <- predict(markCapENet95Model, xTest, s = 0.2595) 
rmseNet95Test <- sqrt(mean((Net95PredTest - yTest)^2)) 
rmseNet95Test


percentChange <- round(((rmseNet95Test/rmseNet95PredTrain)-1)*100, digits = 2)
percentChange
```

R-Squared = 78.21


### More Elastic Net Models, alpha = .975

```{r marketCapNet97.5}
markCapENet975Model <- glmnet(xTrain, yTrain, alpha = .975)
markCapENet975Model
```

```{r marketCapNet97.5Coef}
coef(markCapENet975Model, s = 0.253)
```

```{r marketCapNet975Predict2}
Net975PredTrain <- predict(markCapENet975Model, xTrain, s = 0.253) 
rmseNet975PredTrain <- sqrt(mean((Net975PredTrain - yTrain)^2)) 
rmseNet975PredTrain

Net975PredTest <- predict(markCapENet975Model, xTest, s = 0.253) 
rmseNet975Test <- sqrt(mean((Net975PredTest - yTest)^2)) 
rmseNet975Test


percentChange <- round(((rmseNet975Test/rmseNet975PredTrain)-1)*100, digits = 2)
percentChange
```


```{r marketCapElastic975ResidualPlot}
residual = Net975PredTest - yTest
plot(Net975PredTest, residual)
```

```{r marketCapElastic975ResidualQQPlot}
qqnorm(residual, pch = 1, frame = FALSE)
qqline(residual, col = "steelblue", lwd = 2)
```

```{r marketCapElastic975ResidualStdResidualPlot}
residualStd = residual/sd(residual)
plot(Net975PredTest, residualStd)
```

```{r marketCapElastic975ResidualStdResidualPercent}
x = residualStd[ residualStd > -3 & residualStd < 3]
length(x)/length(residualStd)
```

```{r marketCapElastic975PlotModel}
plot(markCapENet975Model)
```